---
title: "Census Data"
author: "Ally Rakus"
date: "7/16/2020"
output: github_document
---
This is the R markdown file where I work with census data for Baltimore, MD at the census tract level. In this file, I call the data with an API, create variables, and visualize some of the data on an interactive map. 


Calling the appropriate libraries needed for analysis and installing my Census API Key 
```{r}
library(tidyverse)
library(tidycensus)
library(sf)

census_api_key("d432da892f8a6ce576e962e7170b27cf7ef13f8d", overwrite = TRUE, install = TRUE) 

```

Calling the variables I need: income, poverty rate, educational attainment, unemployment, & race

1. Calling for income and poverty rate
```{r}
bmore_census_inc_pov <- get_acs(
  geography = "tract", #tract level
  state = "MD", #in Maryland
  county = c("510"), #Baltimore City
  variables = c(
    "hh_inc_med" = "B19013_001", # median household income
    "pov_pct" = "C17002_001" # poverty rate
  ), 
  survey = "acs5", #5 Year American Community Survey 
  output = "wide", 
  geometry = FALSE)

bmore_census_inc_pov #viewing the new dataframe 
```

2. Calling for the employment status table->allow me to make an unemployment rate variable 
```{r}
bmore_census_unemploy <- get_acs(
  geography = "tract", 
  state = "MD",
  county = c("510"),
  table = "B23001", #employment status table 
  survey = "acs5",
  output = "wide",
  geometry = FALSE)
bmore_census_unemploy
```
Creating variables: labor force & unemployed to then make unemployment rate
```{r}
bmore_census_unemployment <- bmore_census_unemploy %>% 
  rowwise() %>% 
  mutate(
    
    laborforce_est = sum(
      c_across(matches("B23001_0{0,2}(6|13|20|27|34|41|48|55|62|69|74|79|84|92|99|106|113|120|127|134|141|148|155|160|165|170)E"))
    ), #creating labor force estimate
    
    laborforce_moe = moe_sum(
      moe = c_across(matches("B23001_0{0,2}(6|13|20|27|34|41|48|55|62|69|74|79|84|92|99|106|113|120|127|134|141|148|155|160|165|170)M")),
      estimate = laborforce_est
    ), #creating labor force margin of error
    
    pop_unemployed_est = sum(
      c_across(c(matches("B23001_0{0,2}(8|15|22|29|36|43|50|57|64|71|76|81|86|94|101|108|115|122|129|136|143|150|157|162|167|172)E")))
    ), #creating unemployed estimate
    
    pop_unemployed_moe = moe_sum(
      moe = c_across(c(matches("B23001_0{0,2}(8|15|22|29|36|43|50|57|64|71|76|81|86|94|101|108|115|122|129|136|143|150|157|162|167|172)M"))),
      estimate = pop_unemployed_est)
  )%>% 
  ungroup() %>%
  mutate(
    unemploy_rate_est = pop_unemployed_est / na_if(laborforce_est,0), #creating unemployment rate
    
    unemploy_rate_moe = moe_prop(pop_unemployed_est, na_if(laborforce_est,0), pop_unemployed_moe, laborforce_moe) #creating unemployment rate margin of error
  )

# just for this preview, remove all columns matching this pattern, so we can
# more easily see our new columns
bmore_census_unemployment %>% 
  select(-matches("B\\d{5}_\\d{3}."))
```

3. Calling the education table to create educational attainment variables
```{r}
bmore_census_educ <- get_acs(
  geography = "tract", 
  state = "MD",
  county = c("510"),
  table = "B15003", #table with education info
  survey = "acs5",
  output = "wide",
  geometry = FALSE)
bmore_census_educ
```

Creating the variables: no high school education and some college or more education (2 education groups)
```{r}
bmore_census_education <- bmore_census_educ %>%
  mutate(
    pop_edu_denom_est = B15003_001E, #renaming the variables from the table
    pop_edu_denom_moe = B15003_001M,
  ) %>%
  rowwise() %>%
  mutate(
    pop_edu_nohs_num_est = sum(c_across(c(matches("B15003_0{0,2}(2|3|4|5|6|7|8|9|10|11|12|13|14|15|16)E")))), #creating the no high school education variable
    
    pop_edu_nohs_num_moe = moe_sum(
      moe = c_across(c(matches("B15003_0{0,2}(2|3|4|5|6|7|8|9|10|11|12|13|14|15|16)M"))),
      estimate = pop_edu_nohs_num_est
    ),#creating the no HS ed margin of error variable
    
    pop_edu_somecoll_plus_est = sum(c_across(c(matches("B15003_0{0,2}(19|20|21|22|23|24|25)E")))),#creating the some college plus education variable
    
    pop_edu_somecoll_plus_moe = moe_sum(
      moe = c_across(c(matches("B15003_0{0,2}(19|20|21|22|23|24|25)M"))),
      estimate = pop_edu_somecoll_plus_est
    ) #creating the some college plus margin of error variable
  ) %>%
  
#below - creating the percentage variables of each education group
  ungroup() %>% 
  mutate(
    noHS_educ_pct = pop_edu_nohs_num_est / na_if (pop_edu_denom_est, 0),
    noHS_educ_pct_moe = pop_edu_nohs_num_moe / na_if (pop_edu_denom_moe, 0),
    somecollege_plus_pct = pop_edu_somecoll_plus_est / na_if (pop_edu_denom_est, 0),
    somecollege_plus_pct_moe = pop_edu_somecoll_plus_moe / na_if (pop_edu_denom_moe,0)
  )
  
bmore_census_education %>% 
  select(-matches("B\\d{5}_\\d{3}."))
```

Calling the race variables 
```{r}
bmore_census_r<- get_acs(
  geography = "tract", 
  state = "MD",
  county = c("510"),
  variables = c(
    "race_denom" = "B02001_001", #denominator to calculate race percentage
    "black_num" = "B02009_001"), #numerator to calculate the percentage of black residents
  survey = "acs5",
  output = "wide",
  geometry = TRUE) %>% 
  as_tibble() %>% # for nicer printing
  st_as_sf() # make sure to add back the spatial class

bmore_census_r
```

Creating the race variable: percentage of black alone and black and other races in a tract 
```{r}
bmore_census_race <- bmore_census_r %>% 
  mutate(
    pct_black_est = black_numE / na_if(race_denomE,0),
    pct_black_moe = moe_prop(black_numE, na_if(race_denomE,0), black_numM, race_denomM)
  )

bmore_census_race
```

Creating one dataframe for all the variables
```{r}
# Because "bmore_census_race" is a spatial dataframe, when we join onto it all
# the resulting dataframes keep the spatial class.

bmore_census_data_f <- left_join(bmore_census_race, bmore_census_education, by = "GEOID")

bmore_census_data_fi <- left_join(bmore_census_data_f, bmore_census_unemployment, by = "GEOID")

bmore_census_data_merged <- left_join(bmore_census_data_fi, bmore_census_inc_pov, by = "GEOID")

#creating the final dataframe keeping only important variables 
bmore_census_data_MERGED_FINAL <- bmore_census_data_merged %>% select(GEOID, pct_black_est, pct_black_moe, noHS_educ_pct, noHS_educ_pct_moe, somecollege_plus_pct, somecollege_plus_pct_moe, unemploy_rate_est, unemploy_rate_moe, hh_inc_medE, hh_inc_medM, pov_pctE, pov_pctM)

bmore_census_data_MERGED_FINAL
```

Creating a map of census tracts including unemployment rate and percentage of black residents 
```{r}
#calling the needed packages
library(leaflet)
library(htmlwidgets)

#color palette for black pop percentage layer
pop_blck_pct_pal <- colorBin(
  palette = viridisLite::viridis(5), 
  domain = bmore_census_data_MERGED_FINAL$pct_black_est, 
  n = 4
)

#color palette for unemployment rate 
unemployment_pal <- colorBin(
  palette = viridisLite::viridis(5), 
  domain = bmore_census_data_MERGED_FINAL$unemploy_rate_est, 
  n = 4
)

#creating the map
bmore_census_map_race_UE <- bmore_census_data_MERGED_FINAL %>% 
  st_transform(4326) %>% 
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  
#plotting the black pop percentage layer 
  addPolygons(
    fillColor = ~pop_blck_pct_pal(pct_black_est),
    fillOpacity = 0.7,
    color = "white",
    weight = 0.5,
    group = "Black Population Percentage"
  ) %>% 
  
#plotting the unemployment rate layer 
  addPolygons(
    fillColor = ~unemployment_pal(unemploy_rate_est),
    fillOpacity = 0.7,
    color = "white",
    weight = 0.5,
    group = "Unemployment Rate"
  ) %>% 

#plotting the control panel   
  addLayersControl(
    overlayGroups = c("Black Population Percentage", "Unemployment Rate"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% 
  showGroup(c("Black Population Percentage")) %>% 

#creating the legend for black pop percentage
  addLegend(
    position = "topleft", 
    pal = pop_blck_pct_pal, 
    values = ~pct_black_est, 
    title = "Black Population Percentage<br>by Census Tract (2014-2018)",
    labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x)
  ) %>% 

#creating the legend for the unemployment rate    
  addLegend(
    position = "topright", 
    pal = unemployment_pal, 
    values = ~unemploy_rate_est, 
    title = "Unemployment Rate<br>by Census Tract (2014-2018)",
    labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x)
  )

bmore_census_map_race_UE

#saving the map as an html link 
htmlwidgets::saveWidget(bmore_census_map_race_UE, "bmore_census_map_race_UE.html", selfcontained = TRUE)

```

Prepping dataframe of only unemployment data - to be used for correlation analysis: 
```{r}
bmore_unemployment_rate_raw <- bmore_census_data_MERGED_FINAL %>% 
  select(GEOID, unemploy_rate_est, unemploy_rate_moe)

bmore_unemployment_rate_raw
```


