---
title: "Census Data"
author: "Ally Rakus"
date: "6/25/2020"
output: github_document
---

Creating my Census API Key 
```{r}
library(tidyverse)
library(tidycensus)
library(sf)

census_api_key("d432da892f8a6ce576e962e7170b27cf7ef13f8d", overwrite = TRUE, install = TRUE) 

```

Calling the variables I need: income, poverty rate, educational attainment, unemployment, race/ethnicity

1. Calling for income and poverty rate 
```{r}
bmore_census_inc_pov <- get_acs(
  geography = "tract", 
  state = "MD",
  county = c("510"),
  variables = c(
    "hh_inc_med" = "B19013_001", # median household inc
    "pov_pct" = "C17002_001" # poverty rate
  ), 
  survey = "acs5",
  output = "wide",
  geometry = FALSE)
bmore_census_inc_pov
```
2. Calling for the unemployment table
```{r}
bmore_census_unemploy <- get_acs(
  geography = "tract", 
  state = "MD",
  county = c("510"),
  table = "B23001",
  survey = "acs5",
  output = "wide",
  geometry = FALSE)
bmore_census_unemploy
```
Creating variables: labor force, unemployed, unemployment rate
```{r}
bmore_census_unemployment <- bmore_census_unemploy %>% 
  rowwise() %>% 
  mutate(
    
    laborforce_est = sum(
      c_across(matches("B23001_0{0,2}(6|13|20|27|34|41|48|55|62|69|74|79|84|92|99|106|113|120|127|134|141|148|155|160|165|170)E"))
    ),
    
    laborforce_moe = moe_sum(
      moe = c_across(matches("B23001_0{0,2}(6|13|20|27|34|41|48|55|62|69|74|79|84|92|99|106|113|120|127|134|141|148|155|160|165|170)M")),
      estimate = laborforce_est
    ),
    
    pop_unemployed_est = sum(
      c_across(c(matches("B23001_0{0,2}(8|15|22|29|36|43|50|57|64|71|76|81|86|94|101|108|115|122|129|136|143|150|157|162|167|172)E")))
    ),
    
    pop_unemployed_moe = moe_sum(
      moe = c_across(c(matches("B23001_0{0,2}(8|15|22|29|36|43|50|57|64|71|76|81|86|94|101|108|115|122|129|136|143|150|157|162|167|172)M"))),
      estimate = pop_unemployed_est)
  ) %>% 
  ungroup() %>%
  mutate(
    unemploy_rate_est = pop_unemployed_est / na_if(laborforce_est,0),
    unemploy_rate_moe = moe_prop(pop_unemployed_est, na_if(laborforce_est,0), pop_unemployed_moe, laborforce_moe)
  )

# just for this preview, remove all columns matching this pattern, so we can
# more easily see our new columns
bmore_census_unemployment %>% 
  select(-matches("B\\d{5}_\\d{3}."))
```
3. Calling the education table
```{r}
bmore_census_educ <- get_acs(
  geography = "tract", 
  state = "MD",
  county = c("510"),
  table = "B15003",
  survey = "acs5",
  output = "wide",
  geometry = FALSE)
bmore_census_educ
```
Creating variables: no high school education, some college or more education 
```{r}
bmore_census_education <- bmore_census_educ %>%
  mutate(
    pop_edu_denom_est = B15003_001E,
    pop_edu_denom_moe = B15003_001M,
  ) %>%
  rowwise() %>%
  mutate(
    pop_edu_nohs_num_est = sum(c_across(c(matches("B15003_0{0,2}(2|3|4|5|6|7|8|9|10|11|12|13|14|15|16)E")))),
    pop_edu_nohs_num_moe = moe_sum(
      moe = c_across(c(matches("B15003_0{0,2}(2|3|4|5|6|7|8|9|10|11|12|13|14|15|16)M"))),
      estimate = pop_edu_nohs_num_est
    ),
    pop_edu_somecoll_plus_est = sum(c_across(c(matches("B15003_0{0,2}(19|20|21|22|23|24|25)E")))),
    pop_edu_somecoll_plus_moe = moe_sum(
      moe = c_across(c(matches("B15003_0{0,2}(19|20|21|22|23|24|25)M"))),
      estimate = pop_edu_somecoll_plus_est
    )
  ) %>%
  ungroup() %>% 
  mutate(
    noHS_educ_pct = pop_edu_nohs_num_est / na_if (pop_edu_denom_est, 0),
    noHS_educ_pct_moe = pop_edu_nohs_num_moe / na_if (pop_edu_denom_moe, 0),
    somecollege_plus_pct = pop_edu_somecoll_plus_est / na_if (pop_edu_denom_est, 0),
    somecollege_plus_pct_moe = pop_edu_somecoll_plus_moe / na_if (pop_edu_denom_moe,0)
  )
  
bmore_census_education %>% 
  select(-matches("B\\d{5}_\\d{3}."))
```

Calling the race variables 
```{r}
bmore_census_r<- get_acs(
  geography = "tract", 
  state = "MD",
  county = c("510"),
  variables = c(
    "race_denom" = "B02001_001",
    "black_num" = "B02009_001"),
  survey = "acs5",
  output = "wide",
  geometry = TRUE) %>% 
  as_tibble() %>% # for nicer printing
  st_as_sf() # make sure to add back the spatial class

bmore_census_r
```

Creating the race variable: percent black or black and other races 
```{r}
bmore_census_race <- bmore_census_r %>% 
  mutate(
    pct_black_est = black_numE / na_if(race_denomE,0),
    pct_black_moe = moe_prop(black_numE, na_if(race_denomE,0), black_numM, race_denomM)
  )

bmore_census_race
```

Creating one dataframe for all the new variables/dataframes
```{r}
# Because "bmore_census_race" is a spatial dataframe, when we join onto it all
# the resulting dataframes keep the spatial class.

bmore_census_data_f <- left_join(bmore_census_race, bmore_census_education, by = "GEOID")

bmore_census_data_fi <- left_join(bmore_census_data_f, bmore_census_unemployment, by = "GEOID")

bmore_census_data_merged <- left_join(bmore_census_data_fi, bmore_census_inc_pov, by = "GEOID")

bmore_census_data_MERGED_FINAL <- bmore_census_data_merged %>% select(GEOID, pct_black_est, pct_black_moe, noHS_educ_pct, noHS_educ_pct_moe, somecollege_plus_pct, somecollege_plus_pct_moe, unemploy_rate_est, unemploy_rate_moe, hh_inc_medE, hh_inc_medM, pov_pctE, pov_pctM)

bmore_census_data_MERGED_FINAL
```

```{r}
library(leaflet)
library(htmlwidgets)

#color palette for black population layer
pop_blck_pct_pal <- colorBin(
  palette = viridisLite::viridis(5), 
  domain = bmore_census_data_MERGED_FINAL$pct_black_est, 
  n = 4
)
#color palette for unemployment rate 
unemployment_pal <- colorBin(
  palette = viridisLite::viridis(5), 
  domain = bmore_census_data_MERGED_FINAL$unemploy_rate_est, 
  n = 4
)

#creating the map
census_map_final <- bmore_census_data_MERGED_FINAL %>% 
  st_transform(4326) %>% 
  leaflet() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>%
  
#plotting the black population layer 
  addPolygons(
    fillColor = ~pop_blck_pct_pal(pct_black_est),
    fillOpacity = 0.7,
    color = "white",
    weight = 0.5,
    group = "Black Population Percentage"
  ) %>% 
  
#plotting the unemployment rate layer 
  addPolygons(
    fillColor = ~unemployment_pal(unemploy_rate_est),
    fillOpacity = 0.7,
    color = "white",
    weight = 0.5,
    group = "Unemployment Rate"
  ) %>% 

  
  addLayersControl(
    overlayGroups = c("Black Population Percentage", "Unemployment Rate"),
    options = layersControlOptions(collapsed = FALSE)
  ) %>% 
  showGroup(c("Black Population Percentage")) %>% 

#creating the legend for black population
  addLegend(
    position = "topleft", 
    pal = pop_blck_pct_pal, 
    values = ~pct_black_est, 
    title = "Black Population Percentage<br>by Census Tract (2014-2018)",
    labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x)
  ) %>% 

#creating the legend for the unemployment rate    
  addLegend(
    position = "topright", 
    pal = unemployment_pal, 
    values = ~unemploy_rate_est, 
    title = "Unemployment Rate<br>by Census Tract (2014-2018)",
    labFormat = labelFormat(suffix = "%", transform = function(x) 100 * x)
  )

census_map_final

htmlwidgets::saveWidget(census_map, "census_map_final.html", selfcontained = TRUE)

```

Prepping dataframe of only unemployment data: 
```{r}
bmore_unemployment_rate <- bmore_census_data_MERGED_FINAL %>% 
  select(GEOID, unemploy_rate_est, unemploy_rate_moe)

bmore_unemployment_rate
```


